---
title: "NIK Recurence"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: yes
    fig_caption: yes
    fig_crop: no
    number_sections: yes
  html_document:
    toc: yes
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

# NIK and RRPlots

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The libraries

```{r}
library(survival)
library(FRESA.CAD)
library(Biobase)
op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```


```{r}
if (!require("BiocManager", quietly = TRUE))
{
    install.packages("BiocManager")
    BiocManager::install("seventyGeneData")
}
```

```{r}
library(seventyGeneData)
data(vanDeVijver)
class(vanDeVijver)
```

## Getting the clinical data
```{r}
pdata <- pData(vanDeVijver)

```


## ROC Plots
```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
table(pdata$Posnodes)

pander::pander(table(pdata$TTMevent))
pmroc <- plotModels.ROC(cbind(pdata$TTMevent,-pdata$C1used),name="NIK",thr= -0.4) ## Using paper threshold
par(op)

```

## RR Plot Signature correlation
```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

rdata <- cbind(pdata$TTMevent,pdata$C1used)


RRAnalysisCI <- RRPlot(rdata,atRate=c(0.10,0.05),
                     timetoEvent=pdata$RFS,
                     title="Signature: Breast Cancer",
                     ysurvlim=c(0.00,1.0))

par(op)

```


### CI Performance all data

```{r results = "asis"}

pander::pander(t(RRAnalysisCI$keyPoints),caption="Threshold values")
pander::pander(RRAnalysisCI$c.index$cstatCI,caption="C. Index")
pander::pander(t(RRAnalysisCI$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((RRAnalysisCI$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((RRAnalysisCI$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(RRAnalysisCI$thr_atP),caption="Probability Thresholds")
pander::pander(RRAnalysisCI$surdif,caption="Logrank test")

```


## Node positive data


## RR Plot Signature correlation
```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

NodePdata <- subset(pdata,Posnodes=="y")

rdata <- cbind(NodePdata$TTMevent,NodePdata$C1used)


RRAnalysisPos <- RRPlot(rdata,atRate=c(0.90,0.95),
                     timetoEvent=NodePdata$RFS,
                     title="Signature: Breast Cancer",
                     ysurvlim=c(0.00,1.0))

par(op)

```

### CI Performance positive data

```{r results = "asis"}

pander::pander(t(RRAnalysisPos$keyPoints),caption="Threshold values")
pander::pander(RRAnalysisPos$c.index$cstatCI,caption="C. Index")
pander::pander(t(RRAnalysisPos$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((RRAnalysisPos$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((RRAnalysisPos$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(RRAnalysisPos$thr_atP),caption="Probability Thresholds")
pander::pander(RRAnalysisPos$surdif,caption="Logrank test")

```


## Node Negative data


## RR Plot Signature correlation
```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

NodeNdata <- subset(pdata,Posnodes=="n")

rdata <- cbind(NodeNdata$TTMevent,NodeNdata$C1used)


RRAnalysisNeg <- RRPlot(rdata,atRate=c(0.90,0.95),
                     timetoEvent=NodeNdata$RFS,
                     title="Signature: Breast Cancer",
                     ysurvlim=c(0.00,1.0))
par(op)

```

### CI Performance negative data

```{r results = "asis"}

pander::pander(t(RRAnalysisNeg$keyPoints),caption="Threshold values")
pander::pander(RRAnalysisNeg$c.index$cstatCI,caption="C. Index")
pander::pander(t(RRAnalysisNeg$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((RRAnalysisNeg$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((RRAnalysisNeg$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(RRAnalysisNeg$thr_atP),caption="Probability Thresholds")
pander::pander(RRAnalysisNeg$surdif,caption="Logrank test")

```


## RRPlot Cox Model

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
timeinterval <- 5 # Five years

h0 <- sum(pdata$TTMevent & pdata$RFS <= timeinterval)
h0 <- h0/sum((pdata$RFS > timeinterval) | (pdata$TTMevent==1))

mcox <- coxph(Surv(RFS,TTMevent)~C1used,pdata)
pander::pander(summary(mcox)$coefficients)

index <- predict(mcox,pdata)
rdata <- cbind(pdata$TTMevent,ppoisGzero(index,h0))


RRAnalysisCox <- RRPlot(rdata,atRate=c(0.90,0.95),
                     timetoEvent=pdata$RFS,
                     title="NIK: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

par(op)

```
## Expected time to event

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
toinclude <- rdata[,1]==1 
obstiemToEvent <- pdata[,"RFS"]
tmin<-min(obstiemToEvent)
sum(toinclude)

timetoEvent <- meanTimeToEvent(rdata[,2],timeinterval)
tmax<-max(c(obstiemToEvent,timetoEvent))
lmfit <- lm(obstiemToEvent[toinclude]~0+timetoEvent[toinclude])
sm <- summary(lmfit)
pander::pander(sm)

plot(timetoEvent,obstiemToEvent,
     col=1+rdata[,1],
     xlab="Expected time",
     ylab="Observed time",
     main="Expected vs. Observed",
     xlim=c(tmin,tmax),
     ylim=c(tmin,tmax),
     log="xy")
lines(x=c(tmin,tmax),y=lmfit$coefficients*c(tmin,tmax),lty=1,col="blue")
txt <- bquote(paste(R^2 == .(round(sm$r.squared,3))))
text(tmin+0.005*(tmax-tmin),tmax,txt,cex=0.7)
text(tmin+0.015*(tmax-tmin),tmax,sprintf("Slope=%4.3f",sm$coefficients[1]),cex=0.7)
legend("bottomright",legend=c("No Event","Event","Linear fit"),
             pch=c(1,1,-1),
             col=c(1,2,"blue"),
             lty=c(-1,-1,1)
             )

MADerror2 <- mean(abs(timetoEvent[toinclude]-obstiemToEvent[toinclude]))
pander::pander(MADerror2)


```


### Unadjusted Cox Performance

```{r results = "asis"}

pander::pander(t(RRAnalysisCox$keyPoints),caption="Threshold values")
pander::pander(t(RRAnalysisCox$OERatio$estimate),caption="O/E Ratio")
pander::pander(t(RRAnalysisCox$OE95ci),caption="O/E Mean")
pander::pander(t(RRAnalysisCox$OAcum95ci),caption="O/Acum Mean")
pander::pander(RRAnalysisCox$c.index$cstatCI,caption="C. Index")
pander::pander(t(RRAnalysisCox$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((RRAnalysisCox$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((RRAnalysisCox$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(RRAnalysisCox$thr_atP),caption="Probability Thresholds")
pander::pander(RRAnalysisCox$surdif,caption="Logrank test")

```


## RRPlot Cox Adjusted Model  

This time we will include Lymph node status from pathology report and 
Estrogen receptor alpha expression measurement from microarray 

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

mcox <- coxph(Surv(RFS,TTMevent)~C1used*(ESR1 + Posnodes),pdata)
pander::pander(summary(mcox)$coefficients)


index <- predict(mcox,pdata)
rdata <- cbind(pdata$TTMevent,ppoisGzero(index,h0))


RRAnalysisAdCox <- RRPlot(rdata,atRate=c(0.90,0.95),
                     timetoEvent=pdata$RFS,
                     title="Adjusted: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)
par(op)

```


## Expected time to event

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

timetoEvent <- meanTimeToEvent(rdata[,2],timeinterval)
tmax<-max(c(obstiemToEvent,timetoEvent))
lmfit <- lm(obstiemToEvent[toinclude]~0+timetoEvent[toinclude])
sm <- summary(lmfit)
pander::pander(sm)

plot(timetoEvent,obstiemToEvent,
     col=1+rdata[,1],
     xlab="Expected time",
     ylab="Observed time",
     main="Expected vs. Observed",
     xlim=c(tmin,tmax),
     ylim=c(tmin,tmax),
     log="xy")
lines(x=c(tmin,tmax),y=lmfit$coefficients*c(tmin,tmax),lty=1,col="blue")
txt <- bquote(paste(R^2 == .(round(sm$r.squared,3))))
text(tmin+0.005*(tmax-tmin),tmax,txt,cex=0.7)
text(tmin+0.015*(tmax-tmin),tmax,sprintf("Slope=%4.3f",sm$coefficients[1]),cex=0.7)
legend("bottomright",legend=c("No Event","Event","Linear fit"),
             pch=c(1,1,-1),
             col=c(1,2,"blue"),
             lty=c(-1,-1,1)
             )

MADerror2 <-c(MADerror2,mean(abs(timetoEvent-obstiemToEvent)))
pander::pander(MADerror2)

```


### Adjusted Cox Performance

```{r results = "asis"}

pander::pander(t(RRAnalysisAdCox$keyPoints),caption="Threshold values")
pander::pander(t(RRAnalysisAdCox$OERatio$estimate),caption="O/E Ratio")
pander::pander(t(RRAnalysisAdCox$OE95ci),caption="O/E Mean")
pander::pander(t(RRAnalysisAdCox$OAcum95ci),caption="O/Acum Mean")
pander::pander(RRAnalysisAdCox$c.index$cstatCI,caption="C. Index")
pander::pander(t(RRAnalysisAdCox$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((RRAnalysisAdCox$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((RRAnalysisAdCox$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(RRAnalysisAdCox$thr_atP),caption="Probability Thresholds")
pander::pander(RRAnalysisAdCox$surdif,caption="Logrank test")

```



## Calibrating the index

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
calprob <- CoxRiskCalibration(mcox,pdata,"TTMevent","RFS",timeInterval=timeinterval)
pander::pander(c(h0=calprob$h0,
                 Gain=calprob$hazardGain,
                 DeltaTime=calprob$timeInterval),
               caption="Cox Calibration Parameters")

h0 <- calprob$h0
timeinterval <- calprob$timeInterval;

rdata <- cbind(pdata$TTMevent,calprob$prob)


RRAnalysisCalAdCox <- RRPlot(rdata,atRate=c(0.90,0.95),
                     timetoEvent=pdata$RFS,
                     title="Cal. NIK: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

par(op)


```

## Expected time to event

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

timetoEvent <- meanTimeToEvent(rdata[,2],timeinterval)
tmax<-max(c(obstiemToEvent,timetoEvent))
lmfit <- lm(obstiemToEvent[toinclude]~0+timetoEvent[toinclude])
sm <- summary(lmfit)
pander::pander(sm)

plot(timetoEvent,obstiemToEvent,
     col=1+rdata[,1],
     xlab="Expected time",
     ylab="Observed time",
     main="Expected vs. Observed",
     xlim=c(tmin,tmax),
     ylim=c(tmin,tmax),
     log="xy")
lines(x=c(tmin,tmax),y=lmfit$coefficients*c(tmin,tmax),lty=1,col="blue")
txt <- bquote(paste(R^2 == .(round(sm$r.squared,3))))
text(tmin+0.005*(tmax-tmin),tmax,txt,cex=0.7)
text(tmin+0.015*(tmax-tmin),tmax,sprintf("Slope=%4.3f",sm$coefficients[1]),cex=0.7)
legend("bottomright",legend=c("No Event","Event","Linear fit"),
             pch=c(1,1,-1),
             col=c(1,2,"blue"),
             lty=c(-1,-1,1)
             )

MADerror2 <-c(MADerror2,mean(abs(timetoEvent-obstiemToEvent)))
pander::pander(MADerror2)


```


### Calibrated and Adjusted Cox Performance

```{r results = "asis"}

pander::pander(t(RRAnalysisCalAdCox$keyPoints),caption="Threshold values")
pander::pander(t(RRAnalysisCalAdCox$OERatio$estimate),caption="O/E Ratio")
pander::pander(t(RRAnalysisCalAdCox$OE95ci),caption="O/E Mean")
pander::pander(t(RRAnalysisCalAdCox$OAcum95ci),caption="O/Acum Mean")
pander::pander(RRAnalysisCalAdCox$c.index$cstatCI,caption="C. Index")
pander::pander(t(RRAnalysisCalAdCox$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((RRAnalysisCalAdCox$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((RRAnalysisCalAdCox$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(RRAnalysisCalAdCox$thr_atP),caption="Probability Thresholds")
pander::pander(RRAnalysisCalAdCox$surdif,caption="Logrank test")

```

## Comparing Risks

### Comparing concordance Index
```{r results = "asis"}
## Correlation Index
cindex <- RRAnalysisCI$c.index$cstatCI
## Cox Index
cindex <- rbind(cindex,RRAnalysisCox$c.index$cstatCI)
## Adjusted Cox Index
cindex <- rbind(cindex,RRAnalysisAdCox$c.index$cstatCI)
## Adjusted and Calibrated Cox Index
cindex <- rbind(cindex,RRAnalysisCalAdCox$c.index$cstatCI)
rownames(cindex) <- c("CI","Cox","Adj. Cox","Cal. Adj. Cox")

pander::pander(cindex)

```

### Comparing Risk Ratios Index
```{r results = "asis"}
## Correlation Index
RRratio <- c(RR=RRAnalysisCI$keyPoints$RR[1],
             LCI=RRAnalysisCI$keyPoints$RR_LCI[1],
             UCI=RRAnalysisCI$keyPoints$RR_UCI[1])
## Cox Index
RRratio <- rbind(RRratio,c(RR=RRAnalysisCox$keyPoints$RR[1],
                           LCI=RRAnalysisCox$keyPoints$RR_LCI[1],
                           UCI=RRAnalysisCox$keyPoints$RR_UCI[1]))
## Adjusted Cox Index
RRratio <- rbind(RRratio,c(RR=RRAnalysisAdCox$keyPoints$RR[1],
                           LCI=RRAnalysisAdCox$keyPoints$RR_LCI[1],
                           UCI=RRAnalysisAdCox$keyPoints$RR_UCI[1]))
## Adjusted and Calibrated Cox Index
RRratio <- rbind(RRratio,c(RR=RRAnalysisCalAdCox$keyPoints$RR[1],
                           LCI=RRAnalysisCalAdCox$keyPoints$RR_LCI[1],
                           UCI=RRAnalysisCalAdCox$keyPoints$RR_UCI[1]))
rownames(RRratio) <- c("CI","Cox","Adj. Cox","Cal. Adj. Cox")
pander::pander(RRratio)

```


### Comparing logRank values
```{r results = "asis"}
## Correlation Index
SurvDif <- c(chisq=RRAnalysisCI$surdif$chisq,pvalue=RRAnalysisCI$surdif$pvalue)
## Cox Index
SurvDif <- rbind(SurvDif,c(chisq=RRAnalysisCox$surdif$chisq,pvalue=RRAnalysisCox$surdif$pvalue))
## Adjusted Cox Index
SurvDif <- rbind(SurvDif,c(chisq=RRAnalysisAdCox$surdif$chisq,pvalue=RRAnalysisAdCox$surdif$pvalue))
## Adjusted and Calibrated Cox Index
SurvDif <- rbind(SurvDif,c(chisq=RRAnalysisCalAdCox$surdif$chisq,pvalue=RRAnalysisCalAdCox$surdif$pvalue))
rownames(SurvDif) <- c("CI","Cox","Adj. Cox","Cal. Adj. Cox")

pander::pander(SurvDif)

```

### Comparing Sensitivity
```{r results = "asis"}
## Correlation Index
sensi <- RRAnalysisCI$ROCAnalysis$sensitivity
## Cox Index
sensi <- rbind(sensi,RRAnalysisCox$ROCAnalysis$sensitivity)
## Adjusted Cox Index
sensi <- rbind(sensi,RRAnalysisAdCox$ROCAnalysis$sensitivity)
## Adjusted and Calibrated Cox Index
sensi <- rbind(sensi,RRAnalysisCalAdCox$ROCAnalysis$sensitivity)
rownames(sensi) <- c("CI","Cox","Adj. Cox","Cal. Adj. Cox")

pander::pander(sensi)

```

### Comparing Specificity
```{r results = "asis"}
## Correlation Index
speci <- RRAnalysisCI$ROCAnalysis$specificity
## Cox Index
speci <- rbind(speci,RRAnalysisCox$ROCAnalysis$specificity)
## Adjusted Cox Index
speci <- rbind(speci,RRAnalysisAdCox$ROCAnalysis$specificity)
## Adjusted and Calibrated Cox Index
speci <- rbind(speci,RRAnalysisCalAdCox$ROCAnalysis$specificity)
rownames(speci) <- c("CI","Cox","Adj. Cox","Cal. Adj. Cox")
pander::pander(speci)

```


### Comparing O/E
```{r results = "asis"}
OERatio <- NULL
## Cox Index
OERatio <- rbind(OERatio,RRAnalysisCox$OERatio$estimate)
## Adjusted Cox Index
OERatio <- rbind(OERatio,RRAnalysisAdCox$OERatio$estimate)
## Adjusted and Calibrated Cox Index
OERatio <- rbind(OERatio,RRAnalysisCalAdCox$OERatio$estimate)
rownames(OERatio) <- c("Cox","Adj. Cox","Cal. Adj. Cox")

pander::pander(OERatio)

```


### Comparing O/Acum
```{r results = "asis"}
OARatio <- NULL
## Cox Index
OARatio <- rbind(OARatio,RRAnalysisCox$OARatio$estimate)
## Adjusted Cox Index
OARatio <- rbind(OARatio,RRAnalysisAdCox$OARatio$estimate)
## Adjusted and Calibrated Cox Index
OARatio <- rbind(OARatio,RRAnalysisCalAdCox$OARatio$estimate)
rownames(OARatio) <- c("Cox","Adj. Cox","Cal. Adj. Cox")
pander::pander(OARatio)

```

### Comparing NetBenefit
```{r results = "asis"}
NetBen <- NULL
## Cox Index
NetBen <- rbind(NetBen,RRAnalysisCox$keyPoints$NetBenefit)
## Adjusted Cox Index
NetBen <- rbind(NetBen,RRAnalysisAdCox$keyPoints$NetBenefit)
## Adjusted and Calibrated Cox Index
NetBen <- rbind(NetBen,RRAnalysisCalAdCox$keyPoints$NetBenefit)
colnames(NetBen) <- rownames(RRAnalysisCox$keyPoints)
rownames(NetBen) <- c("Cox","Adj. Cox","Cal. Adj. Cox")
pander::pander(NetBen)

```

## Compare the ROC AUC

```{r results = "asis"}

pander::pander(pROC::roc.test(RRAnalysisCI$ROCAnalysis$ROC.analysis$roc.predictor,
                              RRAnalysisAdCox$ROCAnalysis$ROC.analysis$roc.predictor))

```


