---
title: "Risk-Evaluation: Breast Cancer Royston-Altman"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    fig_caption: yes
  pdf_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
    fig_crop: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Evaluation of RISK survival models

This document highlights the use of

-   RRPlot(),

-   CoxRiskCalibration(), and

-   CalibrationProbPoissonRisk(),

for the evaluation (RRPlot), and calibration of cox models (CoxRiskCalibration) or logistic models (CalibrationProbPoissonRisk) of survival data.

Furthermore, it can be used to evaluate any Risk index that reruns the probability of a future event on external data-set.

This document will use the survival::rotterdam, and survival::gbsg data-sets to train and predict the risk of cancer recurrence after surgery. Both Cox and Logistic models will be trained and evaluated.

Here are some sample plots returned by the evaluated functions:

## The libraries

```{r}
library(survival)
library(FRESA.CAD)
op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)
source("~/Documents/GitHub/RISKPLOTS/CODE/auxplots.R")

```

## Breast Cancer Royston-Altman data

### data(gbsg, package="survival") and data(rotterdam, package="survival")

```{r results = "asis"}
gbsgdata <- gbsg
rownames(gbsgdata) <- gbsgdata$pid
gbsgdata$pid <- NULL

odata <-rotterdam
rownames(odata) <- odata$pid
odata$pid <- NULL
odata$rfstime <- odata$rtime
odata$status <- odata$recur
odata$rtime <- NULL
odata$recur <- NULL

odata <- odata[,colnames(odata) %in% colnames(gbsgdata)]

odata$size <- 10*(odata$size=="<=20") + 
  35*(odata$size=="20-50") + 
  60*(odata$size==">50")

data <- as.data.frame(model.matrix(Surv(rfstime,status)~.*age,odata))

data$`(Intercept)` <- NULL

dataBrestCancerTrain <- cbind(time=odata[rownames(data),"rfstime"],status=odata[rownames(data),"status"],data)

colnames(dataBrestCancerTrain) <-str_replace_all(colnames(dataBrestCancerTrain),":","_")
colnames(dataBrestCancerTrain) <-str_replace_all(colnames(dataBrestCancerTrain)," ","")
colnames(dataBrestCancerTrain) <-str_replace_all(colnames(dataBrestCancerTrain),"\\.","_")
colnames(dataBrestCancerTrain) <-str_replace_all(colnames(dataBrestCancerTrain),"-","_")
colnames(dataBrestCancerTrain) <-str_replace_all(colnames(dataBrestCancerTrain),">","_")
dataBrestCancerTrain$time <- dataBrestCancerTrain$time/365 ## To years


pander::pander(table(odata[rownames(data),"status"]),caption="rotterdam")


```

### data(gbsg, package="survival") data conditioning

```{r results = "asis"}

gbsgdata <- gbsgdata[,colnames(odata)]
data <- as.data.frame(model.matrix(Surv(rfstime,status)~.*.,gbsgdata))

data$`(Intercept)` <- NULL

dataBrestCancerTest <- cbind(time=gbsgdata[rownames(data),"rfstime"],status=gbsgdata[rownames(data),"status"],data)

colnames(dataBrestCancerTest) <-str_replace_all(colnames(dataBrestCancerTest),":","_")
colnames(dataBrestCancerTest) <-str_replace_all(colnames(dataBrestCancerTest)," ","")
colnames(dataBrestCancerTest) <-str_replace_all(colnames(dataBrestCancerTest),"\\.","_")
colnames(dataBrestCancerTest) <-str_replace_all(colnames(dataBrestCancerTest),"-","_")
colnames(dataBrestCancerTest) <-str_replace_all(colnames(dataBrestCancerTest),">","_")
dataBrestCancerTest$time <- dataBrestCancerTest$time/365

pander::pander(table(odata[rownames(data),"status"]), caption="gbsg")


```

## Cox Modeling

```{r results = "asis"}

ml <- BSWiMS.model(Surv(time,status)~.,data=dataBrestCancerTrain,loops=1,NumberofRepeats = 5)
sm <- summary(ml)
pander::pander(sm$coefficients)
```

## Cox Model Performance

Here we evaluate the model using the RRPlot() function.

### The evaluation of the raw Cox model with RRPlot()

Here we will use the predicted event probability assuming a baseline hazard for events withing 5 years

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

timeinterval <- 5 # Five years

h0 <- sum(dataBrestCancerTrain$status & dataBrestCancerTrain$time <= timeinterval)
h0 <- h0/sum((dataBrestCancerTrain$time > timeinterval) | (dataBrestCancerTrain$status==1))

pander::pander(t(c(h0=h0,timeinterval=timeinterval)),caption="Initial Parameters")

index <- predict(ml,dataBrestCancerTrain)
rdata <- cbind(dataBrestCancerTrain$status,ppoisGzero(index,h0))

rrAnalysisTrain <- RRPlot(rdata,atRate=c(0.90,0.80),
                     timetoEvent=dataBrestCancerTrain$time,
                     title="Train: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)


```


### By Risk Categories
```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

obsexp <- rrAnalysisTrain$OERatio$atThrEstimates

expObs(obsexp,"Train: Expected vs. Observed")

pander::pander(obsexp)

```



### Time to Event Analysis

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
rrAnalysisdata <- rrAnalysisTrain

pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime,rrAnalysisdata$timetoEventData$expectedTime,paired = TRUE))
highrisk <- rrAnalysisdata$timetoEventData$class == 2
pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime[highrisk],rrAnalysisdata$timetoEventData$expectedTime[highrisk],paired = TRUE))

timesdata <- expObsTime(rrAnalysisdata,title="Train: Expected vs Observed")
pander::pander(timesdata)

```

The Time vs. Events are not calibrated.
Lets do the calibration


### Uncalibrated Performance Report

```{r results = "asis"}

pander::pander(t(rrAnalysisTrain$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisTrain$OERatio$estimate),caption="O/E Ratio")
pander::pander(rrAnalysisTrain$OERatio$atThrEstimates,caption="O/E Ratio")
pander::pander(t(rrAnalysisTrain$OE95ci),caption="O/E Mean")
pander::pander(t(rrAnalysisTrain$OAcum95ci),caption="O/Acum Mean")
pander::pander(rrAnalysisTrain$c.index$cstatCI,caption="C. Index")
pander::pander(t(rrAnalysisTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTrain$thr_atP),caption="Probability Thresholds")
pander::pander(rrAnalysisTrain$surdif,caption="Logrank test")

```

### Cox Calibration

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
op <- par(no.readonly = TRUE)


calprob <- CoxRiskCalibration(ml,dataBrestCancerTrain,"status","time")


pander::pander(c(h0=calprob$h0,
                 Gain=calprob$hazardGain,
                 DeltaTime=calprob$timeInterval),
               caption="Cox Calibration Parameters")

```

### The RRplot() of the calibrated model

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

rcaldata <- cbind(dataBrestCancerTrain$status,calprob$prob)


rrAnalysisCalTrain <- RRPlot(rcaldata,atRate=c(0.90,0.80),
                     timetoEvent=dataBrestCancerTrain$time,
                     title="Cal. Train: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=calprob$timeInterval)

```


### By Risk Categories
```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

obsexp <- rrAnalysisCalTrain$OERatio$atThrEstimates

expObs(obsexp,"Cal. Expected vs. Observed")

pander::pander(obsexp)

```



### Time to Event Analysis

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
rrAnalysisdata <- rrAnalysisCalTrain

pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime,rrAnalysisdata$timetoEventData$expectedTime,paired = TRUE))
highrisk <- rrAnalysisdata$timetoEventData$class == 2
pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime[highrisk],rrAnalysisdata$timetoEventData$expectedTime[highrisk],paired = TRUE))

timesdata <- expObsTime(rrAnalysisdata,title="Train: Expected vs Observed")

pander::pander(timesdata)

```


### Calibrated Train Performance

```{r results = "asis"}

pander::pander(t(rrAnalysisCalTrain$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisCalTrain$OERatio$estimate),caption="O/E Ratio")
pander::pander(t(rrAnalysisCalTrain$OE95ci),caption="O/E Mean")
pander::pander(t(rrAnalysisCalTrain$OAcum95ci),caption="O/Acum Mean")
pander::pander(rrAnalysisCalTrain$c.index$cstatCI,caption="C. Index")
pander::pander(t(rrAnalysisCalTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisCalTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisCalTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisCalTrain$thr_atP),caption="Probability Thresholds")
pander::pander(rrAnalysisCalTrain$surdif,caption="Logrank test")

```

## Performance on the external data set

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

index <- predict(ml,dataBrestCancerTest)
pp <- predictionStats_binary(cbind(dataBrestCancerTest$status,index),plotname="Breast Cancer")
par(op)


prob <- ppoisGzero(index,h0)
rdata <- cbind(dataBrestCancerTest$status,prob)
rrCoxTestAnalysis <- RRPlot(rdata,atThr=rrAnalysisTrain$thr_atP,
                     timetoEvent=dataBrestCancerTest$time,
                     title="Test: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

par(op)

```

### External Data Report

```{r results = "asis"}
pander::pander(t(rrCoxTestAnalysis$keyPoints),caption="Threshold values")
pander::pander(t(rrCoxTestAnalysis$OERatio$estimate),caption="O/E Ratio")
pander::pander(rrCoxTestAnalysis$c.index,caption="C. Index")
pander::pander(t(rrCoxTestAnalysis$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrCoxTestAnalysis$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrCoxTestAnalysis$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrCoxTestAnalysis$thr_atP),caption="Probability Thresholds")
pander::pander(rrCoxTestAnalysis$surdif,caption="Logrank test")

```

### Calibrating the index on the test data

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

calprob <- CoxRiskCalibration(ml,dataBrestCancerTest,"status","time")

pander::pander(c(h0=calprob$h0,
                 Gain=calprob$hazardGain,
                 DeltaTime=calprob$timeInterval),
               caption="Cox Calibration Parameters")

rdata <- cbind(dataBrestCancerTest$status,calprob$prob)

rrAnalysisTest <- RRPlot(rdata,atRate=c(0.90,0.80),
                     timetoEvent=dataBrestCancerTest$time,
                     title="Cal. Test: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=calprob$timeInterval)

```



### By Risk Categories
```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

obsexp <- rrAnalysisTest$OERatio$atThrEstimates

expObs(obsexp,"Cal. Expected vs. Observed")

pander::pander(obsexp)

```



### Time to Event Analysis

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
rrAnalysisdata <- rrAnalysisTest

pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime,rrAnalysisdata$timetoEventData$expectedTime,paired = TRUE))
highrisk <- rrAnalysisdata$timetoEventData$class == 2
pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime[highrisk],rrAnalysisdata$timetoEventData$expectedTime[highrisk],paired = TRUE))

timesdata <- expObsTime(rrAnalysisdata,title="Train: Expected vs Observed")

pander::pander(timesdata)

```


### After Calibration Report

```{r results = "asis"}

pander::pander(t(rrAnalysisTest$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisTest$OERatio$estimate),caption="O/E Ratio")
pander::pander(rrAnalysisTest$c.index,caption="C. Index")
pander::pander(t(rrAnalysisTest$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTest$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTest$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTest$thr_atP),caption="Probability Thresholds")
pander::pander(rrAnalysisTest$surdif,caption="Logrank test")

```

## Logistic Model

Here we train a logistic model on the same data set

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

## Only label subjects that present event withing five years

dataBrestCancerR <- subset(dataBrestCancerTrain, time>=5 | status==1)
dataBrestCancerR$status <- dataBrestCancerR$status * (dataBrestCancerR$time < 5)
dataBrestCancerR$time <- NULL

#ml <- BSWiMS.model(status~1,data=dataBrestCancerR,loops=20,NumberofRepeats = 5)
mlog <- BSWiMS.model(status~1,data=dataBrestCancerR,loops=1,NumberofRepeats = 5)
sm <- summary(mlog)
pander::pander(sm$coefficients)


```

## Logistic Model Performance

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
op <- par(no.readonly = TRUE)


cprob <- predict(mlog,dataBrestCancerTrain)

rdata <- cbind(dataBrestCancerTrain$status,cprob)
rrAnalysisLogTrain <- RRPlot(rdata,atRate=c(0.90,0.80),
                     timetoEvent=dataBrestCancerTrain$time,
                     title="Logistic Train: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=5.0)
par(op)



```


### By Risk Categories
```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

obsexp <- rrAnalysisLogTrain$OERatio$atThrEstimates

expObs(obsexp,"Logistic: Expected vs. Observed")

pander::pander(obsexp)

```



### Time to Event Analysis

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
rrAnalysisdata <- rrAnalysisLogTrain

pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime,rrAnalysisdata$timetoEventData$expectedTime,paired = TRUE))
highrisk <- rrAnalysisdata$timetoEventData$class == 2
pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime[highrisk],rrAnalysisdata$timetoEventData$expectedTime[highrisk],paired = TRUE))

timesdata <- expObsTime(rrAnalysisdata,title="Logistic: Expected vs Observed")

pander::pander(timesdata)

```



### Training Report

```{r results = "asis"}

pander::pander(t(rrAnalysisLogTrain$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisLogTrain$OERatio$estimate),caption="O/E Ratio")
pander::pander(rrAnalysisLogTrain$c.index,caption="C. Index")
pander::pander(t(rrAnalysisLogTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisLogTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisLogTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisLogTrain$thr_atP),caption="Probability Thresholds")
pander::pander(rrAnalysisLogTrain$surdif,caption="Logrank test")

```

### Results on the validation set using Logistic model

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
pre <- predict(mlog,dataBrestCancerTest)
rdata <- cbind(dataBrestCancerTest$status,pre)

rrAnalysisLogTest <- RRPlot(rdata,atThr=rrAnalysisTrain$thr_atP,
                     timetoEvent=dataBrestCancerTest$time,
                     title="Logistic Test: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=5)

par(op)

```


### By Risk Categories
```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

obsexp <- rrAnalysisLogTest$OERatio$atThrEstimates

expObs(obsexp,"Logistic Test: Expected vs. Observed")

pander::pander(obsexp)

```



### Time to Event Analysis

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
rrAnalysisdata <- rrAnalysisLogTest

pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime,rrAnalysisdata$timetoEventData$expectedTime,paired = TRUE))
highrisk <- rrAnalysisdata$timetoEventData$class == 2
pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime[highrisk],rrAnalysisdata$timetoEventData$expectedTime[highrisk],paired = TRUE))

timesdata <- expObsTime(rrAnalysisdata,title="Logistic Test: Expected vs Observed")

pander::pander(timesdata)

```


### Validation Report

```{r results = "asis"}

pander::pander(t(rrAnalysisLogTest$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisLogTest$OERatio$estimate),caption="O/E Ratio")
pander::pander(rrAnalysisLogTest$c.index,caption="C. Index")
pander::pander(t(rrAnalysisLogTest$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisLogTest$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisLogTest$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisLogTest$thr_atP),caption="Probability Thresholds")
pander::pander(rrAnalysisLogTest$surdif,caption="Logrank test")

```


## Logistic Model Poisson Calibration

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

riskdata <- cbind(dataBrestCancerTrain$status,predict(mlog,dataBrestCancerTrain,type="prob"),dataBrestCancerTrain$time)
calprob <- CalibrationProbPoissonRisk(riskdata)

pander::pander(c(h0=calprob$h0,
                 Gain=calprob$hazardGain,
                 DeltaTime=calprob$timeInterval),
               caption="Logistic Calibration Parameters")

timeinterval <- calprob$timeInterval;
gain <- calprob$hazardGain

rdata <- cbind(dataBrestCancerTrain$status,calprob$prob)


rrAnalysisLogCalTrain <- RRPlot(rdata,atRate=c(0.90,0.80),
                     timetoEvent=dataBrestCancerTrain$time,
                     title="Cal. Logistic Train: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)


par(op)


```


### By Risk Categories
```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

obsexp <- rrAnalysisLogCalTrain$OERatio$atThrEstimates

expObs(obsexp,"Logistic Cal: Expected vs. Observed")

pander::pander(obsexp)

```



### Time to Event Analysis

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
rrAnalysisdata <- rrAnalysisLogCalTrain

pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime,rrAnalysisdata$timetoEventData$expectedTime,paired = TRUE))
highrisk <- rrAnalysisdata$timetoEventData$class == 2
pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime[highrisk],rrAnalysisdata$timetoEventData$expectedTime[highrisk],paired = TRUE))

timesdata <- expObsTime(rrAnalysisdata,title="Logistic Cal: Expected vs Observed")

pander::pander(timesdata)

```


### Report of the calibrated logistic: training

```{r results = "asis"}

pander::pander(t(rrAnalysisLogCalTrain$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisLogCalTrain$OERatio$estimate),caption="O/E Ratio")
pander::pander(rrAnalysisLogCalTrain$c.index,caption="C. Index")
pander::pander(t(rrAnalysisLogCalTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisLogCalTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisLogCalTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisLogCalTrain$thr_atP),caption="Probability Thresholds")
pander::pander(rrAnalysisLogCalTrain$surdif,caption="Logrank test")

```

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
probLog <- predict(mlog,dataBrestCancerTest)
aprob <- adjustProb(probLog,gain)

rdata <- cbind(dataBrestCancerTest$status,aprob)
rrAnalysisTestLogistic <- RRPlot(rdata,atThr=rrAnalysisTrain$thr_atP,
                     timetoEvent=dataBrestCancerTest$time,
                     title="Cal. Logistic Test: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)
par(op)

```



### By Risk Categories
```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

obsexp <- rrAnalysisTestLogistic$OERatio$atThrEstimates

expObs(obsexp,"Logistic Test: Expected vs. Observed")

pander::pander(obsexp)

```



### Time to Event Analysis

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
rrAnalysisdata <- rrAnalysisTestLogistic

pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime,rrAnalysisdata$timetoEventData$expectedTime,paired = TRUE))
highrisk <- rrAnalysisdata$timetoEventData$class == 2
pander::pander(wilcox.test(rrAnalysisdata$timetoEventData$eTime[highrisk],rrAnalysisdata$timetoEventData$expectedTime[highrisk],paired = TRUE))

timesdata <- expObsTime(rrAnalysisdata,title="Logistic Test: Expected vs Observed")

pander::pander(timesdata)

```

### Report of the calibrated validation

```{r results = "asis"}

pander::pander(t(rrAnalysisTestLogistic$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisTestLogistic$OERatio$estimate),caption="O/E Ratio")
pander::pander(rrAnalysisTestLogistic$c.index,caption="C. Index")
pander::pander(t(rrAnalysisTestLogistic$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTestLogistic$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTestLogistic$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTestLogistic$thr_atP),caption="Probability Thresholds")
pander::pander(rrAnalysisTestLogistic$surdif,caption="Logrank test")

```

## Comparing the COX and Logistic Models on the Independent Data

```{r results = "asis", warning = FALSE , dpi=600, fig.height= 4.0, fig.width= 8.0}

pander::pander(t(rrCoxTestAnalysis$OAcum95ci))
pander::pander(t(rrAnalysisTestLogistic$OAcum95ci))

pander::pander(t(rrCoxTestAnalysis$OE95ci))
pander::pander(t(rrAnalysisTestLogistic$OE95ci))
maxobs <- sum(dataBrestCancerTest$status)

par(mfrow=c(1,2),cex=0.75)

plot(rrCoxTestAnalysis$CumulativeOvs[,1:2],type="l",lty=1,
     main="Cumulative Probability",
     xlab="Observed",
     ylab="Cumulative Probability",
     ylim=c(0,maxobs),
     xlim=c(0,maxobs))
lines(rrAnalysisTestLogistic$CumulativeOvs[,1:2],lty=2,col="red")
lines(x=c(0,maxobs),y=c(0,maxobs),lty=3,col="gray")
legend("topleft",legend = c("Cox","Logistic","Ideal"),
       col=c("black","red","gray"),
       lty=c(1,2,3),
       cex=0.75
)

dxcox <- rrCoxTestAnalysis$CumulativeOvs$Cumulative-
       rrCoxTestAnalysis$CumulativeOvs$Observed

dxlogit <- rrAnalysisTestLogistic$CumulativeOvs$Cumulative-
       rrAnalysisTestLogistic$CumulativeOvs$Observed

miny <- min(c(dxcox,dxlogit))
maxy <- max(c(dxcox,dxlogit))
plot(rrCoxTestAnalysis$CumulativeOvs$Observed,
     dxcox,
     main="Cumulative Risk Difference",
     xlab="Observed",
     ylab="Cumulative Risk - Observed",
     type="l",
     ylim=c(miny,maxy),
     lty=1)
lines(rrAnalysisTestLogistic$CumulativeOvs$Observed,
     dxlogit,
     lty=2,
     col="red")
legend("topleft",legend = c("Cox","Logistic"),
       col=c("black","red"),
       lty=c(1,2),
       cex=0.75
)

plot(rrCoxTestAnalysis$OEData[,2:3],type="l",lty=1,
     main="Expected over Time",
     xlab="Observed",
     ylab="Expected",
     ylim=c(0,maxobs),
     xlim=c(0,maxobs))
lines(rrAnalysisTestLogistic$OEData[,2:3],lty=2,col="red")
lines(x=c(0,maxobs),y=c(0,maxobs),lty=3,col="gray")
legend("topleft",legend = c("Cox","Logistic","Ideal"),
       col=c("black","red","gray"),
       lty=c(1,2,3),
       cex=0.75
)

coxdif <- rrCoxTestAnalysis$OEData$Expected-
       rrCoxTestAnalysis$OEData$Observed

logdif <- rrAnalysisTestLogistic$OEData$Expected-
       rrAnalysisTestLogistic$OEData$Observed

miny <- min(c(coxdif,logdif))
maxy <- max(c(coxdif,logdif))

plot(rrCoxTestAnalysis$OEData$Observed,
     coxdif,
      ylim=c(miny,maxy),

     main="Expected vs Observed Difference",
     xlab="Observed",
     ylab="Cumulative - Observed",
     type="l",
     lty=1)
lines(rrAnalysisTestLogistic$OEData$Observed,
     logdif,
     lty=2,col="red")

legend("bottomleft",legend = c("Cox","Logistic"),
       col=c("black","red"),
       lty=c(1,2),
       cex=0.75
)

par(op)

```

