---
title: "Breast Cancer: Wisconsin"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(survival)
library(FRESA.CAD)
source("~/GitHub/FRESA.CAD/R/RRPlot.R")
source("~/GitHub/FRESA.CAD/R/PoissonEventRiskCalibration.R")

```

```{r}


dataBreast <- read.csv("~/GitHub/RISKPLOTS/DATA/wpbc.data", header=FALSE)
table(dataBreast$V2)

rownames(dataBreast) <- dataBreast$V1
dataBreast$V1 <- NULL
dataBreast$status <- 1*(dataBreast$V2=="R")
dataBreast$V2 <- NULL
dataBreast$time <- dataBreast$V3
dataBreast$V3 <- NULL
dataBreast <- sapply(dataBreast,as.numeric)
dataBreast <- as.data.frame(dataBreast[complete.cases(dataBreast),])
table(dataBreast$status)
```

## Modeling
```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
ml <- BSWiMS.model(Surv(time,status)~1,data=dataBreast)
sm <- summary(ml)
pander::pander(sm$coefficients)
```

## Cox Model Performance

Here we evaluate the model using the RRPlot() function.

### The evaluation of the raw Cox model with RRPlot()

Here we will use the predicted event probability assuming a baseline hazard for events withing 5 years

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

index <- predict(ml,dataBreast)
timeinterval <- 2*mean(dataBreast$time)

h0 <- sum(dataBreast$status & dataBreast$time < timeinterval)
h0 <- h0/nrow(subset(dataBreast,time > timeinterval | status==1))

rdata <- cbind(dataBreast$status,ppoisGzero(index,h0))

rrAnalysisTrain <- RRPlot(rdata,atProb=c(0.90,0.80),
                     timetoEvent=dataBreast$time,
                     title="Raw Train: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

```

As we can see the Observed probability as well as the Time vs. Events are not calibrated.

### Uncalibrated Performance Report

```{r results = "asis"}

pander::pander(t(rrAnalysisTrain$OERatio),caption="O/E Ratio")
pander::pander(rrAnalysisTrain$c.index,caption="C. Index")
pander::pander(t(rrAnalysisTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTrain$thr_atP),caption="Probability Thresholds")
pander::pander(t(rrAnalysisTrain$RR_atP),caption="Risk Ratio")
pander::pander(rrAnalysisTrain$sufdif,caption="Logrank test")

```



### Cox Calibration

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
op <- par(no.readonly = TRUE)


calprob <- CoxRiskCalibration(ml,dataBreast,"status","time")

pander::pander(c(h0=calprob$h0,
                 Gain=calprob$hazardGain,
                 DeltaTime=calprob$timeInterval),
               caption="Cox Calibration Parameters")

```

### The RRplot() of the calibrated model

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
h0 <- calprob$h0
timeinterval <- calprob$timeInterval;

rdata <- cbind(dataBreast$status,calprob$prob)


rrAnalysisTrain <- RRPlot(rdata,atProb=c(0.90,0.80),
                     timetoEvent=dataBreast$time,
                     title="Train: Breast",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

```

### Calibrated Train Performance

```{r results = "asis"}

pander::pander(t(rrAnalysisTrain$OERatio),caption="O/E Ratio")
pander::pander(rrAnalysisTrain$c.index,caption="C. Index")
pander::pander(t(rrAnalysisTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTrain$thr_atP),caption="Probability Thresholds")
pander::pander(t(rrAnalysisTrain$RR_atP),caption="Risk Ratio")
pander::pander(rrAnalysisTrain$sufdif,caption="Logrank test")

```
