---
title: 'Breast Cancer: Wisconsin'
author: "Jose Tamez"
date: '`r Sys.Date()`'
output:
  pdf_document: 
    toc: yes
    fig_caption: yes
    fig_crop: no
    number_sections: yes
  html_document: 
    toc: yes
editor_options:
  chunk_output_type: console
---
# Wisconsin Prognosis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Libraries

```{r}
library(survival)
library(FRESA.CAD)
op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
#pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)

```

## The data

```{r}


dataBreast <- read.csv("~/GitHub/RISKPLOTS/DATA/wpbc.data", header=FALSE)
table(dataBreast$V2)

rownames(dataBreast) <- dataBreast$V1
dataBreast$V1 <- NULL
dataBreast$status <- 1*(dataBreast$V2=="R")
dataBreast$V2 <- NULL
dataBreast$time <- dataBreast$V3
dataBreast$V3 <- NULL
dataBreast <- sapply(dataBreast,as.numeric)
dataBreast <- as.data.frame(dataBreast[complete.cases(dataBreast),])
table(dataBreast$status)
```

## Modeling
```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
ml <- BSWiMS.model(Surv(time,status)~1,data=dataBreast)
sm <- summary(ml)
pander::pander(sm$coefficients)


```

## Cox Model Performance

Here we evaluate the model using the RRPlot() function.

### The evaluation of the raw Cox model with RRPlot()

Here we will use the predicted event probability assuming a baseline hazard for events withing 5 years

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

index <- predict(ml,dataBreast)
timeinterval <- 2*mean(subset(dataBreast,status==1)$time)

h0 <- sum(dataBreast$status & dataBreast$time <= timeinterval)
h0 <- h0/sum((dataBreast$time > timeinterval) | (dataBreast$status==1))
pander::pander(t(c(h0=h0,timeinterval=timeinterval)),caption="Initial Parameters")

rdata <- cbind(dataBreast$status,ppoisGzero(index,h0))
rownames(rdata) <- rownames(dataBreast)

rrAnalysisTrain <- RRPlot(rdata,atProb=c(0.90),
                     timetoEvent=dataBreast$time,
                     title="Raw Train: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)



```

As we can see the Observed probability as well as the Time vs. Events are not calibrated.

### Uncalibrated Performance Report

```{r results = "asis"}


pander::pander(t(rrAnalysisTrain$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisTrain$OERatio$estimate),caption="O/E Ratio")
pander::pander(t(rrAnalysisTrain$OE95ci),caption="O/E Mean")
pander::pander(t(rrAnalysisTrain$OAcum95ci),caption="O/Acum Mean")
pander::pander(rrAnalysisTrain$c.index$cstatCI,caption="C. Index")
pander::pander(t(rrAnalysisTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTrain$thr_atP),caption="Probability Thresholds")
pander::pander(t(rrAnalysisTrain$RR_atP),caption="Risk Ratio")
pander::pander(rrAnalysisTrain$surdif,caption="Logrank test")

```



### Cox Calibration

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
op <- par(no.readonly = TRUE)


calprob <- CoxRiskCalibration(ml,dataBreast,"status","time")

pander::pander(c(h0=calprob$h0,
                 Gain=calprob$hazardGain,
                 DeltaTime=calprob$timeInterval),
               caption="Cox Calibration Parameters")

```

### The RRplot() of the calibrated model

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
h0 <- calprob$h0
timeinterval <- calprob$timeInterval;

rdata <- cbind(dataBreast$status,calprob$prob)


rrAnalysisTrain <- RRPlot(rdata,atProb=c(0.90),
                     timetoEvent=dataBreast$time,
                     title="Calibrated Train: Breast",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

```

### Calibrated Train Performance

```{r results = "asis"}

pander::pander(t(rrAnalysisTrain$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisTrain$OERatio$estimate),caption="O/E Ratio")
pander::pander(t(rrAnalysisTrain$OE95ci),caption="O/E Mean")
pander::pander(t(rrAnalysisTrain$OAcum95ci),caption="O/Acum Mean")
pander::pander(t(rrAnalysisTrain$c.index$cstatCI),caption="C. Index")
pander::pander(t(rrAnalysisTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTrain$thr_atP),caption="Probability Thresholds")
pander::pander(t(rrAnalysisTrain$RR_atP),caption="Risk Ratio")
pander::pander(rrAnalysisTrain$surdif,caption="Logrank test")

```

## Cross-Validation 

Here we use the estimated h0 and timeinterval from the full set

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

rcv <- randomCV(theData=dataBreast,
                theOutcome = Surv(time,status)~1,
                fittingFunction=BSWiMS.model, 
                trainFraction = 0.9,
                repetitions=100,
                classSamplingType = "Pro"
         )

stp <- rcv$survTestPredictions
stp <- stp[!is.na(stp[,4]),]

bbx <- boxplot(unlist(stp[,1])~rownames(stp),plot=FALSE)
times <- bbx$stats[3,]
status <- boxplot(unlist(stp[,2])~rownames(stp),plot=FALSE)$stats[3,]
prob <- ppoisGzero(boxplot(unlist(stp[,4])~rownames(stp),plot=FALSE)$stats[3,],h0)

rdatacv <- cbind(status,prob)
rownames(rdatacv) <- bbx$names
names(times) <- bbx$names

rrAnalysisTest <- RRPlot(rdatacv,atThr = rrAnalysisTrain$thr_atP,
                     timetoEvent=times,
                     title="Test: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

```

### Cross-Validation Test Performance

```{r results = "asis"}

pander::pander(t(rrAnalysisTest$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisTest$OERatio$estimate),caption="O/E Ratio")
pander::pander(t(rrAnalysisTest$OE95ci),caption="O/E Mean")
pander::pander(t(rrAnalysisTest$OAcum95ci),caption="O/Acum Mean")
pander::pander(rrAnalysisTest$c.index$cstatCI,caption="C. Index")
pander::pander(t(rrAnalysisTest$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTest$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTest$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTest$thr_atP),caption="Probability Thresholds")
pander::pander(t(rrAnalysisTest$RR_atP),caption="Risk Ratio")
pander::pander(rrAnalysisTest$surdif,caption="Logrank test")

```


### Calibrating the test results

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
rdatacv <- cbind(status,prob,times)
calprob <- CalibrationProbPoissonRisk(rdatacv)

pander::pander(c(h0=calprob$h0,
                 Gain=calprob$hazardGain,
                 DeltaTime=calprob$timeInterval),
               caption="Cox Calibration Parameters")

timeinterval <- calprob$timeInterval;

rdata <- cbind(status,calprob$prob)


rrAnalysisTest <- RRPlot(rdata,atProb=c(0.90),
                     timetoEvent=times,
                     title="Calibrated Test: Breast",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

```
### Calibrated Test Performance

```{r results = "asis"}

pander::pander(t(rrAnalysisTest$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisTest$OERatio$estimate),caption="O/E Ratio")
pander::pander(t(rrAnalysisTest$OE95ci),caption="O/E Mean")
pander::pander(t(rrAnalysisTest$OAcum95ci),caption="O/Acum Mean")
pander::pander(rrAnalysisTest$c.index$cstatCI,caption="C. Index")
pander::pander(t(rrAnalysisTest$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTest$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTest$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTest$thr_atP),caption="Probability Thresholds")
pander::pander(t(rrAnalysisTest$RR_atP),caption="Risk Ratio")
pander::pander(rrAnalysisTest$surdif,caption="Logrank test")

```

