---
title: 'Breast Cancer: Wisconsin'
author: "Jose Tamez"
date: '`r Sys.Date()`'
output:
  html_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
  pdf_document: 
    toc: yes
    fig_caption: yes
    fig_crop: no
    number_sections: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Showcasing RRPlots

### Libraries
```{r}
library(survival)
library(FRESA.CAD)
#source("~/GitHub/FRESA.CAD/R/RRPlot.R")
#source("~/GitHub/FRESA.CAD/R/PoissonEventRiskCalibration.R")
op <- par(no.readonly = TRUE)
pander::panderOptions('digits', 3)
#pander::panderOptions('table.split.table', 400)
pander::panderOptions('keep.trailing.zeros',TRUE)
layout(matrix(1:1, nrow=1))
source("C:/Users/jtame/Documents/GitHub/RISKPLOTS/CODE/auxplots.R")

```

### Wisconsin Data Set

```{r}


dataBreast <- read.csv("~/GitHub/RISKPLOTS/DATA/wpbc.data", header=FALSE)
table(dataBreast$V2)

rownames(dataBreast) <- dataBreast$V1
dataBreast$V1 <- NULL
dataBreast$status <- 1*(dataBreast$V2=="R")
dataBreast$V2 <- NULL
dataBreast$time <- dataBreast$V3
dataBreast$V3 <- NULL
dataBreast <- sapply(dataBreast,as.numeric)
dataBreast <- as.data.frame(dataBreast[complete.cases(dataBreast),])
table(dataBreast$status)
```


## Exploring Raw Features with RRPlot

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
convar <- colnames(dataBreast)[lapply(apply(dataBreast,2,unique),length) > 10]
convar <- convar[convar != "time"]
topvar <- univariate_BinEnsemble(dataBreast[,c("status",convar)],"status")
pander::pander(topvar)
topv <- min(5,length(topvar))
topFive <- names(topvar)[1:topv]
RRanalysis <- list();
idx <- 1
topf <- topFive[1]
for (topf in topFive)
{
  RRanalysis[[idx]] <- RRPlot(cbind(dataBreast$status,dataBreast[,topf]),
                              atRate=c(0.90,0.80),
                  timetoEvent=dataBreast$time,
                  title=topf,
#                  plotRR=FALSE
                  )
  idx <- idx + 1
}
names(RRanalysis) <- topFive

```

## Reporting the Metrics

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
pander::pander(RRanalysis[[1]]$keyPoints,caption=topFive[1])
pander::pander(RRanalysis[[2]]$keyPoints,caption=topFive[2])

RRanalysis[[2]]$keyPoints["@MAX_BACC",c("BACC","RR")]

ROCAUC <- NULL
CstatCI <- NULL
LogRangp <- NULL
Sensitivity <- NULL
Specificity <- NULL
MAXBACC <- NULL
RREst <- NULL

for (topf in topFive)
{
  CstatCI <- rbind(CstatCI,RRanalysis[[topf]]$c.index$cstatCI)
  LogRangp <- rbind(LogRangp,RRanalysis[[topf]]$surdif$pvalue)
  Sensitivity <- rbind(Sensitivity,RRanalysis[[topf]]$ROCAnalysis$sensitivity)
  Specificity <- rbind(Specificity,RRanalysis[[topf]]$ROCAnalysis$specificity)
  ROCAUC <- rbind(ROCAUC,RRanalysis[[topf]]$ROCAnalysis$aucs)
  MAXBACC <- rbind(MAXBACC,RRanalysis[[topf]]$keyPoints["@MAX_BACC",c("BACC")])
  RREst <- rbind(RREst,RRanalysis[[topf]]$keyPoints[1,c("RR")])
}
rownames(CstatCI) <- topFive
rownames(LogRangp) <- topFive
rownames(Sensitivity) <- topFive
rownames(Specificity) <- topFive
rownames(ROCAUC) <- topFive
rownames(MAXBACC) <- topFive
rownames(RREst) <- topFive

pander::pander(ROCAUC)
pander::pander(CstatCI)
pander::pander(LogRangp)
pander::pander(Sensitivity)
pander::pander(Specificity)
pander::pander(MAXBACC)
pander::pander(RREst)

meanMatrix <- cbind(ROCAUC[,1],CstatCI[,1],RREst,Sensitivity[,1],Specificity[,1],MAXBACC)
colnames(meanMatrix) <- c("ROCAUC","C-Stat","RR","Sen","Spe","MAX_BACC")
pander::pander(meanMatrix)

```


## Modeling
```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
ml <- BSWiMS.model(Surv(time,status)~1,data=dataBreast,NumberofRepeats = 10)
sm <- summary(ml)
pander::pander(sm$coefficients)


```

## Cox Model Performance

Here we evaluate the model using the RRPlot() function.

### The evaluation of the raw Cox model with RRPlot()

Here we will use the predicted event probability assuming a baseline hazard for events

```{r results = "asis", warning = FALSE , dpi=600, fig.height= 4.5, fig.width= 6.0}

index <- predict(ml,dataBreast)
timeinterval <- round(2*mean(subset(dataBreast,status==1)$time),0)

h0 <- sum(dataBreast$status & dataBreast$time <= timeinterval)
h0 <- h0/sum((dataBreast$time > timeinterval) | (dataBreast$status==1))
pander::pander(t(c(h0=h0,timeinterval=timeinterval)),caption="Initial Parameters")

rdata <- cbind(dataBreast$status,ppoisGzero(index,h0))
rownames(rdata) <- rownames(dataBreast)

rrAnalysisTrain <- RRPlot(rdata,atRate=c(0.90,0.80),
                     timetoEvent=dataBreast$time,
                     title="Raw Train: Breast Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)



```


### Uncalibrated Performance Report

```{r results = "asis"}

pander::pander(t(rrAnalysisTrain$keyPoints),caption="Threshold values")
pander::pander(t(rrAnalysisTrain$OERatio$estimate),caption="O/E Test")
pander::pander(t(rrAnalysisTrain$OE95ci),caption="O/E Mean")
pander::pander(t(rrAnalysisTrain$OARatio$estimate),caption="O/Acum Test")
pander::pander(t(rrAnalysisTrain$OAcum95ci),caption="O/Acum Mean")
pander::pander(t(rrAnalysisTrain$c.index$cstatCI),caption="C. Index")
pander::pander(t(rrAnalysisTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTrain$thr_atP),caption="Probability Thresholds")
pander::pander(rrAnalysisTrain$surdif,caption="Logrank test")

```
