---
title: "free light chain"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(survival)
library(FRESA.CAD)
#library(corrplot)
source("~/GitHub/FRESA.CAD/R/RRPlot.R")


```
# RRPLOTS and flchain

```{r}
odata <- flchain
odata$chapter <- NULL
table(odata$death)
rownames(odata) <- c(1:nrow(odata))
data <- as.data.frame(model.matrix(Surv(futime,death)~.*.,odata))

data$`(Intercept)` <- NULL
table(odata[rownames(data),"death"])

dataFL <- cbind(time=odata[rownames(data),"futime"],status=odata[rownames(data),"death"],data)
dataFL$time <- dataFL$time/365
colnames(dataFL) <-str_replace_all(colnames(dataFL)," ","")
colnames(dataFL) <-str_replace_all(colnames(dataFL),"\\.","_")
colnames(dataFL) <-str_replace_all(colnames(dataFL),":","_")
colnames(dataFL) <-str_replace_all(colnames(dataFL),"-","_")
colnames(dataFL) <-str_replace_all(colnames(dataFL),">","_")

trainsamples <- sample(nrow(dataFL),2000)
dataFLTrain <- dataFL[trainsamples,]
dataFLTest <- dataFL[-trainsamples,]

pander::pander(table(dataFLTrain$status))
pander::pander(table(dataFLTest$status))

```

## Modeling
```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
ml <- BSWiMS.model(Surv(time,status)~1,data=dataFLTrain,loops=1)

sm <- summary(ml)
pander::pander(sm$coefficients)
```


## Cox Model Performance

Here we evaluate the model using the RRPlot() function.

### The evaluation of the raw Cox model with RRPlot()

Here we will use the predicted event probability assuming a baseline hazard for events withing 5 years

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

timeinterval <- mean(subset(dataFLTrain,status==1)$time)

h0 <- sum(dataFLTrain$status & dataFLTrain$time <= timeinterval)
h0 <- h0/sum((dataFLTrain$time > timeinterval) | (dataFLTrain$status==1))

pander::pander(t(c(h0=h0,timeinterval=timeinterval)),caption="Initial Parameters")

index <- predict(ml,dataFLTrain)
rdata <- cbind(dataFLTrain$status,ppoisGzero(index,h0))

rrAnalysisTrain <- RRPlot(rdata,atProb=c(0.90,0.80),
                     timetoEvent=dataFLTrain$time,
                     title="Raw Train: FLC",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

```

As we can see the Observed probability as well as the Time vs. Events are not calibrated.

### Uncalibrated Performance Report

```{r results = "asis"}

pander::pander(t(rrAnalysisTrain$OERatio),caption="O/E Ratio")
pander::pander(t(rrAnalysisTrain$OE95ci),caption="O/E Ratio")
pander::pander(t(rrAnalysisTrain$OAcum95ci),caption="O/Acum Ratio")
pander::pander(t(rrAnalysisTrain$c.index$cstatCI),caption="C. Index")
#pander::pander(rrAnalysisTrain$c.index,caption="C. Index")
pander::pander(t(rrAnalysisTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTrain$thr_atP),caption="Probability Thresholds")
pander::pander(t(rrAnalysisTrain$RR_atP),caption="Risk Ratio")
pander::pander(rrAnalysisTrain$sufdif,caption="Logrank test")

```


### Cox Calibration

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
op <- par(no.readonly = TRUE)


calprob <- CoxRiskCalibration(ml,dataFLTrain,"status","time")

pander::pander(c(h0=calprob$h0,
                 Gain=calprob$hazardGain,
                 DeltaTime=calprob$timeInterval),
               caption="Cox Calibration Parameters")

```

### The RRplot() of the calibrated model

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
h0 <- calprob$h0
timeinterval <- calprob$timeInterval;

rdata <- cbind(dataFLTrain$status,calprob$prob)


rrAnalysisTrain <- RRPlot(rdata,atProb=c(0.90,0.80),
                     timetoEvent=dataFLTrain$time,
                     title="Calibrated Train: FLC",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

```

### Calibrated Train Performance

```{r results = "asis"}

pander::pander(t(rrAnalysisTrain$OERatio),caption="O/E Ratio")
pander::pander(t(rrAnalysisTrain$OE95ci),caption="O/E Ratio")
pander::pander(t(rrAnalysisTrain$OAcum95ci),caption="O/Acum Ratio")
pander::pander(t(rrAnalysisTrain$c.index$cstatCI),caption="C. Index")
#pander::pander(rrAnalysisTrain$c.index,caption="C. Index")
pander::pander(t(rrAnalysisTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTrain$thr_atP),caption="Probability Thresholds")
pander::pander(t(rrAnalysisTrain$RR_atP),caption="Risk Ratio")
pander::pander(rrAnalysisTrain$sufdif,caption="Logrank test")

```


## Performance on the test data set

```{r results = "asis", warning = FALSE, dpi=300, fig.height= 5.0, fig.width= 7.0}

index <- predict(ml,dataFLTest)
pp <- predictionStats_binary(cbind(dataFLTest$status,index),plotname="FLC")
par(op)


prob <- ppoisGzero(index,h0)
rdata <- cbind(dataFLTest$status,prob)
rrAnalysis <- RRPlot(rdata,atThr=rrAnalysisTrain$thr_atP,
                     timetoEvent=dataFLTest$time,
                     title="Test: FLC",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

par(op)

```

### External Data Report

```{r results = "asis"}
pander::pander(t(rrAnalysis$OERatio),caption="O/E Ratio")
pander::pander(t(rrAnalysis$OE95ci),caption="O/E Ratio")
pander::pander(t(rrAnalysis$OAcum95ci),caption="O/Acum Ratio")
pander::pander(t(rrAnalysis$c.index$cstatCI),caption="C. Index")
#pander::pander(rrAnalysis$c.index,caption="C. Index")
pander::pander(t(rrAnalysis$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysis$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysis$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysis$thr_atP),caption="Probability Thresholds")
pander::pander(t(rrAnalysis$RR_atP),caption="Risk Ratio")
pander::pander(rrAnalysis$sufdif,caption="Logrank test")

```

