---
title: "Lung Cancer"
author: "Jose Tamez"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(survival)
library(FRESA.CAD)
#library(corrplot)
source("~/GitHub/FRESA.CAD/R/RRPlot.R")
source("~/GitHub/FRESA.CAD/R/PoissonEventRiskCalibration.R")


```

```{r results = "asis"}


data <- model.frame(Surv(time,status)~.*.,lung,na.action=NULL)
colnames(data) <-str_replace_all(colnames(data),":","_")
colnames(data) <-str_replace_all(colnames(data),"\\.","_")

data$inst <- NULL
data$`Surv(time, status)` <- NULL
dataLung <- cbind(time=lung$time/365,status=lung$status-1,data)
dataLung <- dataLung[complete.cases(dataLung),]
pander::pander(table(dataLung$status))
pander::pander(summary(dataLung$time))
```

## Modeling
```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
ml <- BSWiMS.model(Surv(time,status)~1,data=dataLung,loops=100,NumberofRepeats = 10)
sm <- summary(ml)
pander::pander(sm$coefficients)
```

## Cox Model Performance

Here we evaluate the model using the RRPlot() function.

### The evaluation of the raw Cox model with RRPlot()

Here we will use the predicted event probability assuming a baseline hazard for events withing 5 years

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}

index <- predict(ml,dataLung)
timeinterval <- 1 # One year

h0 <- sum(dataLung$status & dataLung$time < timeinterval)
h0 <- h0/nrow(subset(dataLung,time > timeinterval | status==1))

rdata <- cbind(dataLung$status,ppoisGzero(index,h0))

rrAnalysisTrain <- RRPlot(rdata,atProb=c(0.90,0.80),
                     timetoEvent=dataLung$time,
                     title="Raw Train: Lung Cancer",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

```

As we can see the Observed probability as well as the Time vs. Events are not calibrated.

### Uncalibrated Performance Report

```{r results = "asis"}

pander::pander(t(rrAnalysisTrain$OERatio),caption="O/E Ratio")
pander::pander(rrAnalysisTrain$c.index,caption="C. Index")
pander::pander(t(rrAnalysisTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTrain$thr_atP),caption="Probability Thresholds")
pander::pander(t(rrAnalysisTrain$RR_atP),caption="Risk Ratio")
pander::pander(rrAnalysisTrain$sufdif,caption="Logrank test")

```



### Cox Calibration

```{r results = "asis", warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
op <- par(no.readonly = TRUE)


calprob <- CoxRiskCalibration(ml,dataLung,"status","time")

pander::pander(c(h0=calprob$h0,
                 Gain=calprob$hazardGain,
                 DeltaTime=calprob$timeInterval),
               caption="Cox Calibration Parameters")

```

### The RRplot() of the calibrated model

```{r results = "asis",warning = FALSE , dpi=300, fig.height= 5.0, fig.width= 7.0}
h0 <- calprob$h0
timeinterval <- calprob$timeInterval;

rdata <- cbind(dataLung$status,calprob$prob)


rrAnalysisTrain <- RRPlot(rdata,atProb=c(0.90,0.80),
                     timetoEvent=dataLung$time,
                     title="Train: FLC",
                     ysurvlim=c(0.00,1.0),
                     riskTimeInterval=timeinterval)

```

### Calibrated Train Performance

```{r results = "asis"}

pander::pander(t(rrAnalysisTrain$OERatio),caption="O/E Ratio")
pander::pander(rrAnalysisTrain$c.index,caption="C. Index")
pander::pander(t(rrAnalysisTrain$ROCAnalysis$aucs),caption="ROC AUC")
pander::pander((rrAnalysisTrain$ROCAnalysis$sensitivity),caption="Sensitivity")
pander::pander((rrAnalysisTrain$ROCAnalysis$specificity),caption="Specificity")
pander::pander(t(rrAnalysisTrain$thr_atP),caption="Probability Thresholds")
pander::pander(t(rrAnalysisTrain$RR_atP),caption="Risk Ratio")
pander::pander(rrAnalysisTrain$sufdif,caption="Logrank test")

```
